Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massively Distributed Evolutionary Computation

### 2015-07-30 volunteer-clouds-ieee-ci-si.tex
Minor changes and comments for @JJ for issue #11

\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$
\thanks{Manuscript submitted for review on \today.}%
\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
\thanks{E-mail addresses: jmerelo@ugr.es , mario@tectijuana.edu.mx , pacv@ugr.es}%
\thanks{*Corresponding author.}%
}

@@

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client that can be possibly run on a
browser. As such, it can be run in any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

@@

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for 
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
%other technologies?
% changed - JJ
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using Java
compiled programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. This paper also establishes the desirable features of
any {\em global computing} infrastructure, that can also be applied to
distributed evolutionary computation experiments that rely on
volunteer computing: Scalability, heterogeneity, fault tolerance,
adaptive parallelism, safety, anonymity, hierarchy, ease of use and
reasonable performance. Of all these, probably {\em hierarchy}, that
is, the capability of using preferentially resources in the same
organization could be either disregarded or included in adaptivity,
leaving us with eight desirable properties.

@@

Recent works propose the use of cloud computing services for the components of 
client-server distributed EAs. Cloud resources 
can reduce operational costs by outsourcing both hardware and software maintenance 
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what 
is available in most laboratories, allowing researches to 
scale algorithms at reasonable costs. Researchers have proposed the use of 
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
and DropBox~\cite{DBLP:conf/cec/ArenasGMCRL11}. The use of JavaScript to implement server side components, 
gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
in most Platforms as a Service (PaaS) as many support the nodejs framework. Also many web services
have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
Developers can use the same software tools for both client and server side development.
%Ayuda con la mini conclusion

In the work presented in this paper, JavaScript was used in the full
stack, which has the advantage of using the same code for writing
desktop or command line and browser clients, both based in the EC
library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages
expertise acquired in the library itself. 
%The proposed software has the advantage of full stack JavaScript
%development..
%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
Javascript is also compatible with the functional programming paradigm. According to 
\cite{swanresearch2015}, the benefits of using this kind of paradigm in 
metaheuristics optimization allows an improvement in communicability, 
reproducibility, interoperability, automated assembly, knowledge engineering 
and efficiency.

>>>>

\author{Juan-J.~Merelo*$^1$, Mario Garcia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'a-S\'anchez$^1$
\thanks{Manuscript submitted for review on \today.}%
\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnologico de Tijuana}%
\thanks{E-mail addresses: jmerelo@ugr.es , mario@tectijuana.edu.mx , pacv@ugr.es, pablogarcia@ugr.es}%
\thanks{*Corresponding author.}%
}

@@

There is, then, space for systems that do not have any of those
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. NodIO is first an application that
runs in two tiers, a server and a client that can be possibly run on a
browser. As such, it can be run on any computer that loads the web
page where the client is embedded, including possibly many over which
the scientist has no control.

@@

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for 
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
%other technologies?
% changed - JJ
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using Java
compiled programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. This paper also establishes the desirable features of
any {\em global computing} infrastructure, that can also be applied to
distributed evolutionary computation experiments that rely on
volunteer computing: scalability, heterogeneity, fault tolerance,
adaptive parallelism, safety, anonymity, hierarchy, ease of use and
reasonable performance. Of all these, probably {\em hierarchy}, that
is, the capability of using preferentially resources in the same
organization could be either disregarded or included in adaptivity,
leaving us with eight desirable properties.

@@

Recent works propose the use of cloud computing services for the components of 
client-server distributed EAs. Cloud resources 
can reduce operational costs by outsourcing both hardware and software maintenance 
to the cloud provider. Another advantage is that they enable the provisioning of computing resources beyond what 
is available in most laboratories, allowing researches to 
scale algorithms at reasonable costs. Researchers have proposed the use of 
public clouds such as Amazon EC2 \cite{CloudScale}, Google App Engine\cite{di2013towards} 
and DropBox~\cite{mericloud}. The use of JavaScript to implement server side components, %FERGU: he puesto mejor el de revista en lugar del del CEC
gives software developers certain advantages in the cloud paradigm. JavaScript is a first-class citizen
in most Platforms as a Service (PaaS) as many support the Node.js framework 
\footnote{\url{https://strongloop.com/strongblog/a-comparison-of-node-js-paas-hosting-providers/}}. Also many web services %FERGU: he validado esta frase con un footnote, pero si quieres lo paso a referencias
have native JavaScript libraries and use JSON for data interchange. Another advantage is that 
developers can use the same software tools for both client and server side development. 
Besides the advantages to deal with the same language in the full stack as previously mentioned, 
this also facilitates the interoperability of the exchanged data types.
%Ayuda con la mini conclusion FERGU: mira la última frase de arriba

In the work presented in this paper, JavaScript was used in the full 
stack, which has the advantage of using the same code for writing
desktop or command line and browser clients, both based in the EC
library NodEO \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time when writing code and leverages %FERGU: párrafo en pasado "was used" y presente "saves"
expertise acquired in the library itself. %FERGU: tienes info de LoC usadas para la sección experimental y justificar esta frase?
%The proposed software has the advantage of full stack JavaScript
%development..
%FERGU: añadido el siguiente párrafo y referencia para justificar el uso de JS
Javascript is also compatible with the functional programming paradigm. According to 
\cite{swanresearch2015}, the benefits of using this kind of paradigm in 
metaheuristics optimization allows an improvement in communicability, 
reproducibility, interoperability, automated assembly, knowledge engineering 
and efficiency.
